# Basic class

## microtable class

　Many tools can be used for the bioinformatic analysis, such as QIIME[@Caporaso_QIIME_2010], usearch(https://www.drive5.com/usearch/), 
mothur[@Schloss_Introducing_2009], and RDP(http://rdp.cme.msu.edu/).
Although the format of results may be different from various tools, the main files can be classified into the following parts:
(1) OTU or ASV table, i.e. the species-sample abundance table; 
(2) taxonomy table, the taxonomy assignments information table;
(3) phylogenetic tree;
(4) representative sequences.
(5) Generally, it is useful to create a detailed sample information table to store all the sample information, 
including the environmental data and the missing values (NA).

　The microtable class is the basic class and designed to store the basic data for all the downstream analysis in the microeco package.
At least, the OTU table (i.e. species-sample abundance table) should be provided to create microtable object.
Thus, the microtable class can recognize the sample information table is missing and create a default sample table according to
the sample names of otu_table.
To make the file reading more convenient,
we also build another R package file2meco (https://github.com/ChiLiubio/file2meco) to read the output files of some tools into microtable object.
Currently, those tools/softwares include not only commonly-used QIIME [@Caporaso_QIIME_2010] and QIIME2[@Bolyen_Reproducible_2019], 
but also some metagenomic tools, such as HUMAnN [@Franzosa_Species_2018].
In this tutorial, we use the data inside the package microeco to show some operations.

　The 16S rRNA sequencing results in the example data of the package is used to show the main part of the tutorial.
This dataset is the 16S rRNA gene Miseq sequencing results of wetland soils in China published by An et al.[@An_Soil_2019], 
who surveyed soil bacterial communities in Chinese inland wetlands (IW), 
coastal wetland (CW) and Tibet plateau wetlands (TW) using 16S rRNA gene amplicon sequencing method.
These wetlands include both saline and non-saline samples.
The sample information table have 4 columns: "SampleID", "Group", "Type" and "Saline".
The column "SampleID" is same with the rownames.
The column "Group" represents the IW, CW and TW.
The column "Type" represents the sampling region: northeastern region (NE), northwest region (NW), North China area (NC),
middle-lower reaches of the Yangtze River (YML), southern coastal area (SC), upper reaches of the Yangtze River (YU), Qinghai-Tibet Plateau (QTP). 
The column "Saline" represents the saline soils and non-saline soils.
In this dataset, the environmental factor table is separated from the sample information table.
Another ITS sequencing dataset is also stored in the example data of the package[@Gao_Strong_2019].


```{r, echo = TRUE}
library(microeco)
# load the example data; 16S rRNA gene amplicon sequencing dataset
data(sample_info_16S)
data(otu_table_16S)
data(taxonomy_table_16S)
data(phylo_tree_16S)
# load the environmental data which is detached from sample table
data(env_data_16S)
# use pipe operator in magrittr package
library(magrittr)
# set.seed is used to fix the random number generation to make the results repeatable
set.seed(123)
# make the plotting background same with the tutorial
library(ggplot2)
theme_set(theme_bw())
```

Make sure that the data types of sample_table, otu_table and tax_table are all data.frame as the following part shows.

```{r, echo = TRUE}
class(otu_table_16S)
```

```{r, echo = TRUE, eval = FALSE}
otu_table_16S[1:5, 1:5]
```

```{r, echo = FALSE}
pander::pander(otu_table_16S[1:5, 1:5])
```

```{r, echo = TRUE}
class(taxonomy_table_16S)
```

```{r, echo = TRUE, eval = FALSE}
taxonomy_table_16S[1:5, 1:3]
```

```{r, echo = FALSE}
pander::pander(taxonomy_table_16S[1:5, 1:3])
```

Sometimes, your taxonomic table may have some chaotic information, such NA, unidentified and unknown.
These information can influence the following taxonomic abundance calculation.
So it is necessary to clean this file using the following code.
Another important part of this operation is to unify the taxonomic prefix,
e.g. transforming D_1__ to p__ for phylum level.


```{r, echo = TRUE, eval = FALSE}
# make the taxonomic information unified, important
taxonomy_table_16S %<>% tidy_taxonomy
```


The rownames of sample_table in microtable object (i.e. sample names) are used for selecting samples/groups in all the related operations in the package.
Before you create microtable object, make sure that the rownames of sample information table are the correct sample names.


```{r, echo = TRUE}
class(sample_info_16S)
```

```{r, echo = TRUE, eval = FALSE}
sample_info_16S[1:5, ]
```

```{r, echo = FALSE}
pander::pander(sample_info_16S[1:5, ])
```

In this example, the environmental data is stored in the env_data_16S alone. 
The user can also directly integrate those data into the sample information table.

```{r, echo = TRUE}
class(env_data_16S)
```

```{r, echo = FALSE}
pander::pander(env_data_16S[1:5, 1:5])
```

```{r, echo = TRUE}
class(phylo_tree_16S)
```

Then, we create an object of microtable class.
This operation is very similar with the package phyloseq[@Mcmurdie_phyloseq_2013], but microeco is more brief and simpler.
The otu_table in the microtable class must be the species-sample format: rownames must be OTU names, colnames must be sample names.
The required sample names must be same in rownames of sample_table and colnames of otu_table.

```{r, echo = TRUE}
# In R6 class, '$new' is the original method used to create a new object of class
dataset <- microtable$new(otu_table = otu_table_16S)
# If you only provide abundance table, the class can help you create a sample info table
class(dataset)
dataset
# Let's create a dataset microtable with more information
dataset <- microtable$new(sample_table = sample_info_16S, otu_table = otu_table_16S, tax_table = taxonomy_table_16S, phylo_tree = phylo_tree_16S)
dataset
```

To make the species and sample information consistent across different files in the dataset object, we can use function tidy_dataset() to trim the dataset.


```{r, echo = TRUE}
dataset$tidy_dataset()
print(dataset)
```


Then, we remove OTUs which are not assigned in the Kingdom "k__Archaea" or "k__Bacteria". 
```{r, echo = TRUE}
dataset$tax_table %<>% base::subset(Kingdom == "k__Archaea" | Kingdom == "k__Bacteria")
print(dataset)
```

We also remove OTUs with the taxonomic assignments "mitochondria" or "chloroplast".

```{r, echo = TRUE}
# This will remove the lines containing the taxa word regardless of taxonomic ranks and ignoring word case in the tax_table.
# So if you want to filter some taxa not considerd pollutions, please use subset like the previous operation.
dataset$filter_pollution(taxa = c("mitochondria", "chloroplast"))
print(dataset)
```

Then, to make the OTUs same in otu_table, tax_table and phylo_tree, we use tidy_dataset() again.

```{r, echo = TRUE}
dataset$tidy_dataset()
print(dataset)
```


Then we use sample_sums() to check the sequence numbers in each sample.

```{r, echo = TRUE}
dataset$sample_sums() %>% range
```

Sometimes, in order to reduce the effects of sequencing depth on the diversity measurements, 
we need to perform the resampling to make the sequence number equal for each sample.
The function rarefy_samples can invoke the function tidy_dataset automatically before and after the rarefying.

```{r, echo = TRUE}
# As an example, we use 10000 sequences in each sample
dataset$rarefy_samples(sample.size = 10000)
dataset$sample_sums() %>% range
```

Then, we calculate the taxa abundance at each taxonomic rank using cal_abund().
This function return a list called taxa_abund containing several data frame of the abundance information at each taxonomic rank.
The list is stored in the microtable object automatically.
It's worth noting that the cal_abund() function can be used to solve some complex cases,
such as supporting both the relative and absolute abundance calculation and selecting the partial taxonomic columns.
Those have been shown in README of file2meco package (https://github.com/ChiLiubio/file2meco).

```{r, echo = TRUE}
dataset$cal_abund()
# return dataset$taxa_abund
class(dataset$taxa_abund)
```

The function save_abund() can be used to save the taxa abundance file to a local place.

```{r, echo = TRUE, eval = FALSE}
dir.create("taxa_abund")
dataset$save_abund(dirpath = "taxa_abund")
```

Then, let's calculate the alpha diversity.
The result is also stored in the object microtable automatically.
For the definition of each alpha diversity index, please see http://scikit-bio.org/docs/latest/generated/skbio.diversity.alpha.html
As an example, we do not calculate phylogenetic diversity.

```{r, echo = TRUE}
# If you want to add Faith's phylogenetic diversity, use PD = TRUE, this will be a little slow
dataset$cal_alphadiv(PD = FALSE)
# return dataset$alpha_diversity
class(dataset$alpha_diversity)
```

```{r, echo = TRUE, eval = FALSE}
# save dataset$alpha_diversity to a directory
dir.create("alpha_diversity")
dataset$save_alphadiv(dirpath = "alpha_diversity")
```

We also calculate the distance matrix of beta diversity using function cal_betadiv().
We provide four most frequently used indexes: Bray-curtis, Jaccard, weighted Unifrac and unweighted unifrac.

```{r, echo = FALSE, eval = TRUE, message = FALSE}
invisible(dataset$cal_betadiv(unifrac = FALSE))
```

```{r, echo = TRUE, eval = FALSE}
# If you do not want to calculate unifrac metrics, use unifrac = FALSE
# require GUniFrac package
dataset$cal_betadiv(unifrac = TRUE)
# return dataset$beta_diversity
class(dataset$beta_diversity)
# save dataset$beta_diversity to a directory
dir.create("beta_diversity")
dataset$save_betadiv(dirpath = "beta_diversity")
```




