# Pipeline


Pipelining operations (i.e., method chaining) can make the steps of data analysis more concise and clear. 
The microeco package has fully supported pipelining operations since version 1.12.0. 
Based on the characteristics of R6 class objects, pipelining operations can be performed using the `$` operator. 
In this section, we list some examples.


```{r, echo = TRUE, eval = FALSE}
# version >= 1.12.0
if(packageVersion("microeco") < "1.12.0"){
	install.packages("microeco")
}
```

When you need to name an object of a class, 
do not include plotting function in pipeline operations, because plotting functions return the plot object (usually ggplot2), not the class object itself.

```{r, echo = TRUE, eval = FALSE}
library(magrittr)
library(microeco)

t1 <- trans_alpha$
	new(dataset = dataset, group = "Type")$
	cal_diff(method = "anova", measure = "Shannon")

t1$plot_alpha(plot_type = "errorbar", measure = "Shannon", add_sig = TRUE)
```


When the main purpose is to create a plot and there is no need to retain the object of a class, you can write the entire operation chain.


```{r, echo = TRUE, eval = FALSE}
trans_alpha$
	new(dataset = dataset, group = "Type")$
	cal_diff(method = "anova", measure = "Shannon")$
	plot_alpha(plot_type = "errorbar", measure = "Shannon", add_sig = TRUE)


g1 <- trans_alpha$
	new(dataset = dataset, group = "Type")$
	cal_diff(method = "anova", measure = "Shannon")$
	plot_alpha(plot_type = "errorbar", measure = "Shannon", add_sig = TRUE)

g1
```


Other examples for reference are shown below.

```{r, echo = TRUE, eval = FALSE}
# microtable

library(magrittr)
library(microeco)
data(sample_info_16S)
data(otu_table_16S)
data(taxonomy_table_16S)
data(phylo_tree_16S)
set.seed(123)

test <- microtable$
	new(sample_table = sample_info_16S, otu_table = otu_table_16S, tax_table = taxonomy_table_16S, phylo_tree = phylo_tree_16S)$
	filter_pollution(taxa = c("mitochondria", "chloroplast"))$
	tidy_dataset()$
	filter_taxa()

test$sample_sums() %>% range

test$
	rarefy_samples()$
	cal_abund()$
	save_abund(file.path(test_dir, "taxa_abund"))$
	cal_alphadiv()$
	save_alphadiv(file.path(test_dir, "alpha_diversity"))$
	cal_betadiv()$
	save_betadiv(file.path(test_dir, "beta_diversity"))



# trans_beta

library(magrittr)
library(microeco)
data(dataset)

t1 <- trans_beta$
	new(dataset = dataset, group = "Group", measure = "bray")$
	cal_ordination(method = "PCoA", ncomp = 5)

g1 <- t1$plot_ordination(plot_color = "Group")
g1


trans_beta$
	new(dataset = dataset, group = "Type", measure = "bray")$
	cal_group_distance(within_group = TRUE)$
	cal_group_distance_diff(method = "wilcox")$
	plot_group_distance()

trans_beta$
	new(dataset = dataset, measure = "bray")$
	cal_manova(manova_all = FALSE, group = "Type", by_group = "Group")$
	res_manova

trans_beta$
	new(dataset = dataset, group = "Group", measure = "bray")$
	cal_anosim()$
	res_anosim



# trans_network

library(microeco)
data(dataset)

t1 <- trans_network$new(dataset = dataset, cor_method = "pearson", filter_thres = 0.001)$
	cal_network(COR_p_thres = 0.05, COR_cut = 0.6)$
	cal_module()$
	cal_network_attr()$
	get_node_table()$
	get_edge_table()$
	cal_sum_links()$
	save_network()

t1$plot_network()
t1$plot_sum_links(method = "circlize", transparency = 0.2, annotationTrackHeight = circlize::mm_h(c(5, 5)))
t1$plot_taxa_roles(use_type = 1)



# trans_classifier

library(magrittr)
library(microeco)
data(dataset)

t1 <- trans_classifier$
	new(dataset = dataset, y.response = "Saline", x.predictors = "Genus")$
	cal_preProcess(method = c("center", "scale", "nzv"))$
	cal_feature_sel(boruta.maxRuns = 300, boruta.pValue = 0.01)$
	cal_split(prop.train = 3/4)$
	set_trainControl()$
	cal_train()$
	cal_predict()$
	cal_ROC()

t1$plot_ROC()


t1$
	cal_caretList(methodList = c('rf', 'svmRadial'))$
	cal_caretList_resamples()$
	plot_caretList_resamples() + geom_point(position = "jitter") + theme_bw()


# trans_env

library(microeco)
data(dataset)
data(env_data_16S)

trans_env$
	new(dataset = dataset, add_data = env_data_16S[, 4:10])$
	cal_diff(group = "Group", method = "KW_dunn")$
	plot_diff(measure = "pH", add_sig = T)


t1 <- trans_env$
	new(dataset = dataset, add_data = env_data_16S[, 4:11])$
	cal_ordination(method = "dbRDA", use_measure = "bray")$
	trans_ordination(adjust_arrow_length = TRUE, max_perc_env = 1.5)$
	cal_ordination_anova()

t1$plot_ordination(plot_color = "Group")


trans_env$
	new(dataset = dataset, add_data = env_data_16S[, 4:11])$
	cal_cor(use_data = "Genus")$
	plot_cor()


# trans_func

library(magrittr)
library(microeco)
data(dataset)

trans_func$
	new(dataset)$
	cal_spe_func(prok_database = "FAPROTAX")$
	cal_spe_func_perc(abundance_weighted = FALSE)$
	plot_spe_func_perc()


```
















